---
title: "Main Text Figures and Supplimental Materials for *Temperature alone is not enough: food-web context determines evolutionary responses to warming* "
author: "Ze-Yi Han"
date: "2024-12-04"
output:
  html_document:
    fig_caption: yes
    highlight: pygments
    theme: spacelab
    toc: yes
    toc_depth: 3
    code_folding: show
---

```{=html}
<style>
h1, h2, h3, h4, h5, h6 {
    font-family: "Bookman", "Gerogia";
    color: ##69b3a2
}

a, a:hover {
    color: ##b24e00;
}

pre {
    font-size: 12px;
    color: ##546399;
}
</style>
```

------------------------------------------------------------------------

```{r setup, include=FALSE}
### Loadings----
knitr::opts_chunk$set(eval = TRUE, cache = TRUE, warning=FALSE, message=FALSE, fig.align = "center")
pacman::p_load(doBy,lme4, readr,ggpubr,ggiraph,MuMIn,car,ggplot2, data.table, nlme,tidyverse,moments)
require(dplyr)

```

# data read in

```{r}
### data read in
c0 <- fread("/Users/zeyihan/Documents/PhD/Projects/Temp_Pred_Evo/02_Clean_data/Chlamy_data_submission.csv") %>% dplyr::select(-V1)
c_d0 <- fread("/Users/zeyihan/Documents/PhD/Projects/Temp_Pred_Evo/02_Clean_data/Chlamy_day0_data_submission.csv")%>% dplyr::select(-V1)
pd <- fread("/Users/zeyihan/Documents/PhD/Projects/Temp_Pred_Evo/02_Clean_data/Protist_density.csv")

```

Color platelets

```{r}
strain_col <- c( "#f4ca76","#add0af")  
strain_name <- c("VFL1-1", "WT")

temp_col <- c("#79bdf9","#fedb8a","#f3ac7f")
temp_col2 <- c("#6aa7db","#e7b74a","#e09466")

temp_name <- c(expression(paste("19 ",degree,"C")),expression(paste("22 ",degree,"C")), expression(paste("25 ",degree,"C")))


fig.theme <- theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"), strip.background =element_rect(colour="white", fill="#f0f0f0"),plot.title = element_text(color="black", size=14, face="bold",hjust=0.5),axis.ticks.length=unit(-0.15, "cm"), axis.text.x = element_text(margin=unit(c(0.5,0.5,0.5,0.5), "cm")), axis.text.y = element_text(margin=unit(c(0.5,0.5,0.5,0.5), "cm")))
```

There are a few jars that have Chlmay population crashed from the begining (rep 1 for TetraP.cau treatment in 19**°**C and 20**°**C, jar 25 and 91). Those jars can't offer us helpful information, thus we excluded them from future stats analysis. We replace NA for those jars!

Here we calculate the z-score of each chlamy's size and remove the one's have z-score above 2, so we work with the 95% of our phenotypic data.

```{r results='hide'}

# Set factors classes and levels, log FSC-H/-A and SSC-H/-A

c0_2 <-bind_rows(c_d0,c0) %>%filter(!Jar %in% c(25,91)) %>% mutate(Day=as.numeric(Day),Day_f=as.factor(Day), Temp=as.numeric(Temp), Temp_f=as.factor(Temp),Jar=as.numeric(Jar),Strain = factor(Strain),Trt=factor(Trt,levels=c("Control","Tetra","Glauc","P.cau","TetraP.cau","GlaucP.cau")),PredComp = factor(PredComp,levels=c("Control","No","Yes")), Rep = as.character(Rep),Strain_day=paste(Strain,Day,sep="_"), Temp_Day=paste(Temp,Day,sep="_")) %>% mutate(lgFSC_H=log(FSC_H))

df <- c0_2 %>% group_by(Day,Strain,Temp,Trt) %>% dplyr::mutate( mlgFSC_H = mean(lgFSC_H), sdlgFSC_H = sd(lgFSC_H), z_lgFSC_H = (lgFSC_H -mlgFSC_H)/sdlgFSC_H)
# remove data points with z-score > 2 
c0_trim_fscH <- df %>%filter(z_lgFSC_H > -2 & z_lgFSC_H<2) 
length(c0_trim_fscH$FSC_H)/length(df$FSC_H)
```

Our algae counts are based on 30ul sample. Here convert to ind/ml density. To quantify strain frequency, we first calculated the density of the algal population and density of each chlamy strain, then divide the strain density by the total density.

```{r}
# Calculate total count for each jar
cden <- c0_trim_fscH %>% group_by(Day,Day_f,Jar,Temp,Temp_f,Trt,Rep,PredComp,Temp_Day,Block) %>% dplyr::summarise(Count_tr=length(FSC_H)) %>%
  mutate(Count_tr= ifelse(Day == 0, 4000*0.03,Count_tr), 
         Dens=Count_tr*1000/30, lnDens = log(Dens+1))

# summary by strain in each replicate
cdenSt0 <- c0_trim_fscH %>% group_by(Day,Day_f,Jar,Temp,Temp_f,Trt,Rep,Strain,PredComp,Strain_day,Temp_Day,Block) %>% dplyr::summarise(Count=length(FSC_H)) %>% 
  mutate(Count= ifelse(Day == 0, 2000*0.03,Count),
         Dens=Count*1000/30,lnDens = log(Dens+1))

```

We then calculate the allele frequency of the two strain in each replicate.

```{r results="hide"}
# allele Frequency of VFL1-1
vfl_sum <- filter(cdenSt0,Strain == "VFL1-1") %>% left_join(dplyr::select(cden,Day:Count_tr)) %>% mutate(Freq = Count/Count_tr)
  
# allele Frequency of WT-mNG
WT_sum <- cdenSt0 %>%filter(Strain == "WT-mNG") %>% left_join(dplyr::select(cden,Day:Count_tr)) %>% mutate(Freq = Count/Count_tr)

# check Frequency add up to 1
vfl_sum$Freq+WT_sum$Freq
```

Here is the *Chlamydomonas reinhardtii* data summary of count, density and frequency of each strain in each replicate/jar.

```{r results="hide"}
cdenSt <- rbind(vfl_sum,WT_sum) 
```

We now calculate the mean count, density and frequency of each strain across replicates within each treatment

```{r results="hide"}
# mean density,count and frequency of each treatment (across replication)
cdenStm <-  cdenSt %>% group_by(Day,Day_f,Temp,Temp_f,PredComp,Trt,Strain,Strain_day) %>%
summarise(mCount=mean(Count,na.rm=TRUE),sdCount=sd(Count,na.rm=TRUE),mlnDens=mean(lnDens,na.rm=TRUE),mDens = mean(Dens,na.rm=TRUE),sdDens=sd(Dens,na.rm=TRUE),mFreq=mean(Freq,na.rm=TRUE),sdFreq=sd(Freq,na.rm=TRUE))
# strain mean within treatment
vfl_m <- cdenStm %>% dplyr::filter(Strain == "VFL1-1")
WT_m <- cdenStm %>% dplyr::filter(Strain == "WT-mNG")

# dataframe with No Competition
ncp<- cdenSt %>% dplyr::filter(!PredComp == "Yes") %>% mutate(Strain = factor(Strain, levels=c("WT-mNG","VFL1-1")))
```

# Main Text Figures

## Fig 2

Data from model prediction. Model equations and parameters see supplemental material.

```{r}
ctrl_mp <- read.csv("~/Documents/PhD/Projects/Temp_Pred_Evo/03_Analysis/Math_model/control_final.csv",header = FALSE)  %>% mutate(Trt="Control") 
tetra_mp <- read.csv("~/Documents/PhD/Projects/Temp_Pred_Evo/03_Analysis/Math_model/tetra_final.csv",header = FALSE)%>% mutate(Trt="Tetra")
glauc_mp <- read.csv("~/Documents/PhD/Projects/Temp_Pred_Evo/03_Analysis/Math_model/glauc_final.csv",header = FALSE) %>% mutate(Trt="Glauc")
pcaud_mp <- read.csv("~/Documents/PhD/Projects/Temp_Pred_Evo/03_Analysis/Math_model/pcaud_final.csv",header = FALSE) %>% mutate(Trt="P.cau")

cName <- c("t","WT-mNG","VFL1-1","Predator","WT-freq","VFL-freq","Trt")
colnames(ctrl_mp) <- cName
colnames(tetra_mp) <- cName
colnames(glauc_mp) <- cName
colnames(pcaud_mp) <- cName

modpred <- rbind(ctrl_mp,tetra_mp,glauc_mp,pcaud_mp)
modpred_dens <- modpred %>% dplyr::select(-c(Predator, `WT-freq`,`VFL-freq`)) %>% pivot_longer('WT-mNG':'VFL1-1',names_to = "Strain", values_to = "Dens")
modpred_freq <- modpred %>% dplyr::select(-c(Predator, `WT-mNG`,`VFL1-1`)) %>% pivot_longer('WT-freq':'VFL-freq',names_to = "Strain", values_to = "Freq") %>% mutate(Strain=case_when(Strain=="WT-freq" ~ "WT-mNG",Strain=="VFL-freq"~"VFL1-1"))

modpred2 <- full_join(modpred_dens,modpred_freq) %>% mutate(Strain=factor(Strain, levels=c("WT-mNG","VFL1-1")),Trt=factor(Trt, levels=c("Control", "Tetra","Glauc","P.cau")))
```

Fig 2.a Model prediction of prey clonal density

```{r fig.width = 8.5, fig.height = 2.2}
# model prediction for Fig 2
fig2_a <- ggplot(data=modpred2,aes(x=t, y=Dens, group=Strain,fill=Strain,color=Strain)) + 
  labs(x="",y="Density")+
    scale_color_manual( values= c("#add0af","#f4ca76")  )+
    geom_point(size = 1, alpha =0.5)+
    geom_line()+
    scale_x_continuous(breaks=c( 0,5,10,15),limits = c(0, 15))+
    facet_wrap(~Trt, nrow = 1)+
  fig.theme
fig2_a
```

Fig 2.b Model prediction of prey clonal frequency

```{r fig.width = 8.5, fig.height = 2.2}
fig2_b <- ggplot(data=modpred2,aes(x=t, y=Freq, group=Strain,fill=Strain,color=Strain)) + 
   labs(x = "Time",y="Genetic Frequency", size = 2)+
   geom_bar(position="stack", stat="identity")+
    facet_wrap(~Trt, nrow = 1)+
    xlab("Day")+
    scale_fill_manual(values=c("#add0af","#f4ca76"))+
    scale_color_manual( values=c("#add0af","#f4ca76"))+
    fig.theme
fig2_b
```

Fig 2.c Prey clonal log density from experiment

```{r fig.width = 8.5, fig.height = 2.2}
fig2_c <- ggplot(data=ncp,aes(x=Day, y=Dens, group=Strain,fill=Strain,color=Strain)) + 
  labs(x="",y="Density")+
    scale_color_manual( values= c("#add0af","#f4ca76")  )+
    geom_point(position=position_jitterdodge(jitter.width = 1.5, dodge.width = 4),size = 1, alpha =0.5)+
    stat_summary(fun.y = "mean", geom = "line",position = position_dodge(4),size=0.6)+
     stat_summary(fun.y = "mean", geom = "point",position = position_dodge(4),size = 1)+
    stat_summary(fun.data = mean_se, geom = "errorbar",position = position_dodge(4),width=3, size= 0.6 )+
    facet_wrap(~Trt, nrow = 1)+
  fig.theme
fig2_c
```

Fig 2.d Prey clonal frequency from experiment

```{r fig.width = 8.5, fig.height = 2.2}
fig2_d <- ggplot(data=ncp,aes(x=Day, y=Freq, group=Strain,fill=Strain,color=Strain)) + 
   labs(x = "Day",y="Genetic Frequency", size = 2)+
    stat_summary(fun.y = "mean", geom = "bar",position="stack")+
    facet_wrap(~Trt, nrow = 1)+
    xlab("Day")+
    scale_fill_manual(values=c("#add0af","#f4ca76"))+
    scale_color_manual( values=c("#add0af","#f4ca76"))+
    fig.theme
fig2_d
```

Fig 2.e Prey size for both clones

```{r fig.width=8, fig.height=2.5}
fscH_ncp <- c0_trim_fscH %>% filter(!Trt %in% c("TetraP.cau", "GlaucP.cau"))
fig2_e <- ggplot(fscH_ncp,aes(x=Day, y=lgFSC_H, group=Strain_day,fill = Strain,color=Strain)) + 
    geom_violin(width = 5) +
    stat_summary(fun.y = "mean", geom = "point",position = position_dodge(5), size=1.5, stroke = 1,shape = 1, color="#9c97a8")+
    stat_summary(fun.y = "mean", geom = "point",position = position_dodge(5), size=1.3, shape = 16, color="white")+
    facet_grid(~Trt) +
    labs(y = "log Size")+
    scale_fill_manual(name = "Strain",values=strain_col) + 
    scale_color_manual(name = "Strain",values=strain_col) +
    fig.theme
fig2_e
```

## Fig 3

```{r}
# Single predators treatment data frame for each prey strain
vfl_sng <- vfl_sum %>% filter(PredComp != "Yes", Day!=0)
wt_sng <- WT_sum %>% filter(PredComp != "Yes", Day!=0)
# Competition treatments data frame for each prey strain
vfl_cmp <- vfl_sum %>% filter(!Trt %in% c("Control","P.cau"), Day!=0) %>% ungroup %>% mutate(FocalSp = ifelse(Trt %in%c("Tetra","TetraP.cau"), "Tetra","Glauc" ))
wt_cmp <- WT_sum %>% filter(!Trt %in% c("Control","P.cau"), Day!=0) %>% ungroup %>% mutate(FocalSp = ifelse(Trt %in%c("Tetra","TetraP.cau"), "Tetra","Glauc" ))
```

Here we calculated the changes in vfl1-1 frequencies across temperatures (delta vfl1-1 frequency) by deducting the mean vfl1-1 frequencies at 19C from each replicate at 22C and 25C in each treatment.

```{r}
v19_mean_sng <- vfl_sng %>% filter(Temp == 19 ) %>%  group_by(Temp,Trt,Day) %>% summarize(m19Freq=mean(Freq)) %>% ungroup %>% dplyr::select(-Temp)
vdltf_sng <- vfl_sng %>% filter( !Temp == 19 ) %>% left_join(.,v19_mean_sng) %>% mutate(dlt_Freq_T=Freq-m19Freq)
```

Fig 3a. vfl1-1 delta frequency

```{r fig.width=8.5,fig.height=3}
fig_3a <- ggplot() + 
   stat_summary(vdltf_sng, mapping=aes(x=Day, y=dlt_Freq_T, fill = Temp_f),fun.y = "mean", geom = "bar",position=position_dodge(4), stat="identity",size = 0.6, width = 3)+
   stat_summary(vdltf_sng, mapping=aes(x=Day, y=dlt_Freq_T, color = Temp_f),fun.data = mean_se, geom = "errorbar",position = position_dodge(4),width=2 , size= 0.45) +
  geom_hline(yintercept = 0,color= temp_col[1], size = 1)+ 
    facet_wrap(~Trt,nrow=1)+
    labs(x="Day", y = "Delta VFL1-1 Frequency")+
    scale_fill_manual(name = "Temperature",values=temp_col[2:3])+ 
    scale_color_manual(name = "Temperature",values=temp_col[2:3])+ 
   scale_x_discrete(limits = c(5,10,15))+
   fig.theme
fig_3a
```

Fig 3b. P. caud dens + vfl1-1 freq

```{r}
# add day 0 data for all protists
tetra_d0<- pd %>% filter(Day=="5" & Species=="Tetra") %>% mutate(Day=0,Day_f="0", Dens=5)
glauc_d0<- pd %>% filter(Day=="5" & Species=="Glauc") %>% mutate(Day=0,Day_f="0", Dens=5) 
p.cau_d0<- pd %>% filter(Day=="5" & Species=="P.cau") %>% mutate(Day=0,Day_f="0", Dens=0.5) 
pd_all <- rbind(tetra_d0,glauc_d0,p.cau_d0,pd)
```

Chlmaydomonas population crashed from the beginning in jar 25 and 91, thus we excluded them from future stats analysis.

```{r}
# protists density
pd2 <- pd_all %>% na.omit%>% filter(!Jar %in% c(25,91)) %>% mutate(Day=as.numeric(Day),Day_f=as.factor(Day_f), Temp=as.numeric(Temp), Temp_f=as.factor(Temp_f),Trt=factor(Trt,levels=c("Tetra","Glauc","P.cau","TetraP.cau","GlaucP.cau")),Species=factor(Species,levels=c("Tetra","Glauc","P.cau")), Rep = as.character(Rep), Sp_temp = paste(Species,Temp,sep="_"), SpTempTD = paste(Species,Temp,Day,Trt,sep="_"), lnDens = log(Dens+1))  

pd_sng <- pd2 %>% filter(Trt%in%c("Tetra","Glauc","P.cau") & !Day==0) %>% ungroup %>% dplyr::select(Jar,Day,Temp,Species,Dens,lnDens) %>% mutate(Temp_f=as.factor(Temp))
colnames(pd_sng)[4:6] <- c("PtSpecies","PtDens","PtlnDens")

PtVfl_sng <- left_join(vfl_sng,pd_sng)
PtWt_sng <- left_join(wt_sng,pd_sng)
PtVfl_pcaud <- PtVfl_sng %>% filter(Trt=="P.cau")
shape_pcaud_temp <- PtVfl_pcaud  %>% ungroup %>% mutate(temp_shape = case_when(Temp=="19" ~ 1, Temp=="22"~2, Temp=="25"~0))  %>% pull(temp_shape)
```

```{r fig.width = 5, fig.height = 4.2}
fig_3b <- ggplot(PtVfl_pcaud) + 
#vfl1-1 freq
   stat_summary(mapping=aes(x=Day, y=Freq*8.5, fill=Temp_f),fun.y = "mean", geom = "bar",position=position_dodge(4),alpha = 0.5, width = 4)+
    scale_fill_manual(name = "Temperature",labels = temp_name,values=temp_col) +
    scale_color_manual(name = "Temperature",labels = temp_name,values=temp_col2) +
    scale_shape_manual(name ="Temperature", labels = temp_name,values=c(16,17,15))+
# Protist
  geom_point(aes(x=Day, y=PtlnDens,group =Temp_f,fill=Temp_f,color=Temp_f,shape=Temp_f),position=position_jitterdodge(jitter.width = 1.5, dodge.width = 4),size = 1.5 ,shape=shape_pcaud_temp, alpha=0.7)+
  stat_summary(aes(x=Day, y=PtlnDens,group=Temp_f,color=Temp_f), fun.y = "mean", geom = "line",position = position_dodge(4),size=0.6)+
   stat_summary(aes(x=Day, y=PtlnDens,group=Temp_f,color=Temp_f),fun.data = mean_se, geom = "errorbar",position = position_dodge(4), size= 0.45,width = 3) +
   stat_summary(aes(x=Day, y=PtlnDens,group=Temp_f,color=Temp_f, shape = Temp_f),fun.y = "mean", geom = "point",position = position_dodge(4),size = 2)+
  scale_y_continuous(name = "Protists Ln(Density)", limits = c(0,6.1), sec.axis = sec_axis( trans=~./8.5, name="VFL1-1 Frequency")) +
  fig.theme
fig_3b
```

Fig 3c. wt plasticity by temp in P. caud

```{r  fig.width = 5, fig.height = 2.5}
ws_pcaud <- c0_trim_fscH %>% filter(Strain == "WT-mNG", Trt=="P.cau", Day!=0)

fig_3c <- ggplot(data=ws_pcaud,aes(x=Day, y=lgFSC_H, group=Temp_Day,fill=Temp_f,color=Temp_f)) + 
   geom_violin()+ 
    scale_color_manual(name = "Temperature",labels = temp_name,values=temp_col) + 
    scale_fill_manual(name = "Temperature",labels = temp_name,values=temp_col)+
    ylab("WT Size")+
  stat_summary(fun.y = "mean", geom = "point",position = position_dodge(4.5), size=1.5, stroke = 1,shape = 1, color="#9c97a8")+
  stat_summary(fun.y = "mean", geom = "point",position = position_dodge(4.5), size=1.3, shape = 16, color="white")+
    facet_grid(~Trt) +
  fig.theme
fig_3c
```

## Fig 4

Fig 4

```{r fig.width = 7, fig.height = 6}
tPcau <- vfl_sum %>% filter(Trt %in%c("Tetra","P.cau")) %>% mutate(Trt_Plot = "TetraP.cau")
gPcau <- vfl_sum %>% filter(Trt %in%c("Glauc","P.cau")) %>% mutate(Trt_Plot = "GlaucP.cau")
la <- vfl_sum %>% filter(Trt %in%c("Control","TetraP.cau", "GlaucP.cau")) %>% mutate(Trt_Plot = Trt) 
bar_cdenSt_VFL <-  bind_rows(la,tPcau,gPcau)

# Chlamy with only ComPetition treatment for overlay both predator onto competition(c22cp)
overlay_top <- bar_cdenSt_VFL %>% filter(Trt_Plot%in%c("TetraP.cau", "GlaucP.cau") & Trt %in% c("Tetra", "Glauc","P.cau")) 
overlay_bottom <- bar_cdenSt_VFL %>%filter(Trt%in%c("Control","TetraP.cau", "GlaucP.cau"))
overlay_top_sum <- overlay_top %>% group_by(Trt_Plot, Trt, Temp_f,Day)%>% dplyr::summarise(mFreq = mean(Freq)) 

require(ggpattern)
  fig_4 <- ggplot() + 
    geom_bar_pattern(overlay_top_sum,mapping=aes(Day, mFreq,pattern = Trt ), color = "#b8b2b3",fill  = 'white',pattern_fill = "#b8b2b3", pattern_color  = '#b8b2b3', pattern_spacing = rep(c(0.08, 0.02,0.15, 0.02), each = 4, times=3), pattern_density = rep(c(0.35, 0.2,0.15, 0.2), each = 4,times=3), stat='identity',position = 'dodge') +
    stat_summary(overlay_bottom, mapping=aes(x=Day, y=Freq, fill=Trt),fun.y = "mean", geom = "bar",position=position_dodge(4), stat="identity", alpha = 0.5)+
    facet_grid(Temp_f~factor(Trt_Plot,levels=c("Control","TetraP.cau","GlaucP.cau")))+
    labs(x="Day", y = "VFL1-1 Frequency")+
    scale_pattern_manual(values = c(Tetra = "circle", Glauc = "stripe", P.cau = "circle")) +
    scale_fill_manual(name = "Trt",values=c("#8d8d8d", "#a7a2e8" ,"#9fc359") )+ 
  fig.theme
  fig_4
```

# Appendix II figures

## fig 1

```{r fig.width=10, fig.height=6}
Trt.labs <- c("Control", "Tetrahymena", "Glaucoma","Paramecium","Tetra+Param","Glauc+Param")
names(Trt.labs) <- unique(cdenSt$Trt)

ap.all.eco <- ggplot(data=cdenSt,aes(x=Day, y=Dens,group=Strain,fill=Strain,color=Strain)) + 
    ylab("C. reinhardtii Density") +
    scale_color_manual(name = "Strain",values=strain_col)+
    geom_point(position=position_jitterdodge(jitter.width = 1.5, dodge.width = 4),size = 1, alpha =0.5)+
    stat_summary(fun.y = "mean", geom = "line",position = position_dodge(4),size=0.6)+
    stat_summary(fun.y = "mean", geom = "point",position = position_dodge(4),size = 1)+
    stat_summary(fun.data = mean_se, geom = "errorbar",position = position_dodge(4),width=3, size= 0.6 )+
    facet_grid(Temp~Trt, labeller = labeller(Trt = Trt.labs)) +
  fig.theme;ap.all.eco
```

## fig 2

```{r fig.width=10, fig.height=6}
ap.all.evo <- ggplot(cdenSt, aes(x=Day, y=Freq, fill=Strain)) + 
     stat_summary( aes(x=Day, y=Freq, fill=Strain),fun.y = "mean", geom = "bar",position="stack", stat="identity", alpha=0.9)+
    ylab("Frequency")+
    facet_grid(Temp~Trt, labeller = labeller(Trt = Trt.labs)) +
    scale_fill_manual(name = "Strain",values=strain_col)+
    fig.theme;ap.all.evo
```

## fig 3

```{r fig.width=10, fig.height=6}
AP_all_size <- ggplot(c0_trim_fscH,aes(x=Day, y=lgFSC_H, group=Temp_Day)) + 
  geom_violin(aes(fill = Temp_f,color=Temp_f),trim=TRUE) +
  stat_summary(fun.y = "mean", geom = "point",position = position_dodge(4.5), size=1, stroke = 1,shape = 1, color="#9c97a8")+
  stat_summary(fun.y = "mean", geom = "point",position = position_dodge(4.5), size=0.9, shape = 16, color="white")+
    ylab("Size")+
    facet_grid(Strain~Trt, labeller = labeller(Trt = Trt.labs)) +
    ggtitle("C.reinhardtii Size") +
    scale_fill_manual(name = "Strain",values=temp_col) + 
    scale_color_manual(name = "Strain",values=temp_col) +
  fig.theme;AP_all_size
```

## fig 4

```{r fig.width=18,fig.height=7}
ap_clumping <- ggplot(c0_trim_fscH,aes(x=lgFSC_H, y=lgFSC_A, group=Temp_Day,color=Strain))+
  geom_segment(aes(x=13,xend=16,y=12, yend=15),color = "black",size=0.3)+
  geom_point(size=0.0005, alpha = 0.1)+
  facet_grid(Day+Strain~Trt+Temp)+
   scale_color_manual(name="Strains",values= c("#f49834","#add0af"))+
  theme_bw()+theme(legend.position="right",panel.border = element_blank(), panel.grid.major = element_blank(),panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"), strip.background =element_rect(colour="white", fill="#f0f0f0"),plot.title = element_text(color="black", size=14, face="bold",hjust=0.5),axis.ticks.length=unit(-0.15, "cm"), axis.text.x = element_text(margin=unit(c(0.5,0.5,0.5,0.5), "cm")), axis.text.y = element_text(margin=unit(c(0.5,0.5,0.5,0.5), "cm")),strip.text.x = element_text(size = 14))
ap_clumping
```

## fig 5

```{r fig.width=10, fig.height=10}
# calculate initial population growth rate for each strain of Chlamydomonas
c_r <- cdenSt %>% filter(Day==5, !Dens==0) %>% ungroup %>% dplyr::select(Day_f, Temp, Temp_f,Trt, Rep,Strain,PredComp, Dens) %>% mutate(D0 = case_when(Strain == "VFL1-1" ~ 2000, Strain== "WT-mNG" ~ 2000), r = log(Dens/D0)/5)

# calculate maximum abundance for each strain of Chlamydomonas
c_maxMean <- cdenStm %>% group_by(Temp,Strain,Trt,PredComp) %>% arrange(desc(mlnDens)) %>% slice(1) %>% mutate(StrTempTD = paste(Strain,Temp,Day,Trt,sep="_"))
cdenSt$StrTempTD <- mutate(cdenSt,StrTempTD = paste(Strain,Temp,Day,Trt,sep="_")) %>% pull(StrTempTD)
c_max <- cdenSt %>% filter(StrTempTD %in% c_maxMean$StrTempTD) 

c_r_shape <- c_r %>% ungroup %>% mutate(point_shape = case_when(Temp == 19 ~ 1,Temp  == 22 ~ 2, Temp == 25~0)) %>% orderBy(~ Trt+Temp,.) %>% pull(point_shape)

c_r_fig<- ggplot(c_r, aes(x=Temp_f, y=r, group = Trt, color = Temp_f,shape = Temp_f))+
    scale_color_manual(name = "Temperature",values=temp_col) + 
    scale_shape_manual(name = "Temperature",values=c(16,17,15))+
  xlab("Temperature")+
    geom_point(position=position_jitterdodge(jitter.width = 1.5),size = 1, shape = c_r_shape)+
    stat_summary(fun.y = "mean", geom = "point",size = 2.5,alpha = 0.7)+
    stat_summary(fun.data = mean_se, geom = "errorbar",width=1,size= 0.45) +
  facet_grid(Strain ~ Trt,  scales = "fixed",labeller = labeller(Trt = Trt.labs))+
  fig.theme
c_r_fig

c_maxMean <- cdenStm %>% group_by(Temp,Strain,Trt,PredComp) %>% arrange(desc(mlnDens)) %>% slice(1) %>% mutate(StrTempTD = paste(Strain,Temp,Day,Trt,sep="_"))

cdenSt$StrTempTD <- mutate(cdenSt,StrTempTD = paste(Strain,Temp,Day,Trt,sep="_")) %>% pull(StrTempTD)
c_max <- cdenSt %>%filter(StrTempTD %in% c_maxMean$StrTempTD) 
cmax_point_shape2 <-c_max  %>% ungroup %>% mutate(point_shape = case_when(Temp == 19 ~ 1,Temp  == 22 ~ 2, Temp == 25~0)) %>% orderBy(~ + Strain+ Trt+Day+Block+Temp,.) %>% pull(point_shape)

c_max_fig <- ggplot(c_max , aes(x=Temp_f, y=lnDens, group = Trt, color = Temp_f,shape = Temp_f))+
    scale_color_manual(name = "Temperature",values=temp_col) + 
    scale_shape_manual(name = "Temperature",values=c(16,17,15))+
  ylab("Max Ln Density")+
  xlab("Temperature")+
    geom_point(position=position_jitterdodge(jitter.width = 1.5),size = 1, shape = cmax_point_shape2)+
    #stat_summary(position=position_dodge(2),fun.y = "mean", geom = "line",size=0.6, color ="grey")+
    stat_summary(fun.y = "mean", geom = "point",size = 2.5,alpha = 0.7)+
    stat_summary(fun.data = mean_se, geom = "errorbar",width=1,size= 0.45) +
  facet_grid(Strain~Trt)+
  theme_bw()+theme(legend.position="bottom",panel.border = element_blank(), panel.grid.major = element_blank(),panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"), strip.background =element_rect(colour="white", fill="#f0f0f0"),plot.title = element_text(color="black", size=14, face="bold",hjust=0.5),axis.ticks.length=unit(-0.15, "cm"), axis.text.x = element_text(margin=unit(c(0.5,0.5,0.5,0.5), "cm")), axis.text.y = element_text(margin=unit(c(0.5,0.5,0.5,0.5), "cm")),strip.text.x = element_text(size = 14))

c_r_maxDens <- ggarrange(c_r_fig, c_max_fig,labels = c("A", "B"),ncol = 1, nrow = 2,common.legend = TRUE, legend = "bottom")
c_r_maxDens
```

```{r fig.width=10, fig.height=10}
cdenSt %>% filter(Day==5, !Dens==0) %>% ungroup %>% dplyr::select(Day_f, Temp,Trt,Strain,PredComp, Dens) %>% mutate(D0 = case_when(Strain == "VFL1-1" ~ 2000, Strain== "WT-mNG" ~ 2000), r = log(Dens/D0)/5) %>% group_by(Trt,Strain) %>% dplyr::summarise(mr = mean(r) )

```

## fig 6

```{r fig.width=10, fig.height=8}
pd_r <- pd2 %>% filter(Day==5, !Dens==0) %>% mutate(D0 = case_when(Species == "Tetra" ~ 5,Species == "Glauc" ~ 5, Species == "P.cau"~0.5), r = log(Dens/D0)/5)  

point_shape <- pd_r %>% ungroup %>% mutate(point_shape = case_when(Temp == 19 ~ 1,Temp  == 22 ~ 2, Temp == 25~0)) %>% orderBy(~ + Species+ Trt+Day+Block+Temp,.) %>% pull(point_shape)

g1 <- ggplot(pd_r, aes(x=Temp_f, y=r, group = Trt, color = Temp_f,shape = Temp_f))+
    scale_color_manual(name = "Temperature",values=temp_col) + 
    scale_shape_manual(name = "Temperature",values=c(16,17,15))+
  xlab("Temperature")+
    geom_point(position=position_jitterdodge(jitter.width = 1.5),size = 1, shape = point_shape)+
    stat_summary(fun.y = "mean", geom = "point",size = 2.5,alpha = 0.7)+
    stat_summary(fun.data = mean_se, geom = "errorbar",width=1,size= 0.45) +
  facet_wrap(Species~Trt, nrow=1, scales = "fixed",labeller = labeller(Trt = Trt.labs))+
  fig.theme

pd_sum <- pd2 %>% group_by(Day,Day_f,Block,Temp,Temp_f,Trt,Species,PredComp) %>% dplyr::summarise(mDens=mean(Dens), mlnDens=mean(lnDens))
pd_maxMean <- pd_sum %>% group_by(Temp,Species,PredComp,Trt) %>% dplyr::summarise(max=max(mlnDens))
pd_maxdens <- pd_sum[which(pd_sum$mlnDens %in% pd_maxMean$max),] %>% mutate(SpTempTD = paste(Species,Temp,Day,Trt,sep="_"))
pd_max <- pd2 %>%  filter(SpTempTD %in% pd_maxdens$SpTempTD) 
point_shape2 <- pd_max %>% ungroup %>% mutate(point_shape = case_when(Temp == 19 ~ 1,Temp  == 22 ~ 2, Temp == 25~0)) %>% orderBy(~ + Species+ Trt+Day+Block+Temp,.) %>% pull(point_shape)

g2 <- ggplot(pd_max, aes(x=Temp_f, y=lnDens, group = Trt, color = Temp_f,shape = Temp_f))+
    scale_color_manual(name = "Species",values=temp_col) + 
    scale_shape_manual(name = "Species",values=c(16,17,15))+
  ylab("Max Log Density")+
  xlab("Temperature")+
    geom_point(position=position_jitterdodge(jitter.width = 1.5),size = 1, shape = point_shape2)+
    stat_summary(fun.y = "mean", geom = "point",size = 2.5,alpha = 0.7)+
    stat_summary(fun.data = mean_se, geom = "errorbar",width=1,size= 0.45) +
  facet_wrap(Species~Trt, nrow=1, scales = "fixed",labeller = labeller(Trt = Trt.labs))+
  fig.theme

g3 <- ggarrange(g1, g2,labels = c("a", "b"),ncol = 1, nrow = 2,common.legend = TRUE, legend = "bottom")
g3
```

## fig 7

```{r fig.width = 7, fig.height = 9}
pcomp <- pd2 %>% filter(PredComp=="Yes")  %>% orderBy(~ + Trt+Day+Species+Temp,.)

tPcau <- vfl_sum %>% filter(Trt %in%c("Tetra","P.cau")) %>% mutate(Trt_Plot = "TetraP.cau")
gPcau <- vfl_sum %>% filter(Trt %in%c("Glauc","P.cau")) %>% mutate(Trt_Plot = "GlaucP.cau")
la <- vfl_sum %>% filter(Trt %in%c("Control","TetraP.cau", "GlaucP.cau")) %>% mutate(Trt_Plot = Trt) 
bar_cdenSt_VFL <-  bind_rows(la,tPcau,gPcau)

# Chlamy with only ComPetition treatment for overlay both predator onto competition(c22cp)
overlay_top <- bar_cdenSt_VFL %>% filter(Trt_Plot%in%c("TetraP.cau", "GlaucP.cau") & Trt %in% c("Tetra", "Glauc","P.cau")) 
overlay_bottom <- bar_cdenSt_VFL %>%filter(Trt%in%c("Control","TetraP.cau", "GlaucP.cau"))
overlay_bottom_cmp <- overlay_bottom %>% filter(!Trt=="Control", !Day ==0)

pcomp_pcau <- pcomp %>% filter(Species=="P.cau", !Day ==0) 
pcomp_tetra_glauc <- pcomp %>% filter(!Species=="P.cau", !Day ==0) 
sp_shape_tg <- pcomp_tetra_glauc %>% ungroup %>% mutate(temp_shape = case_when(Temp=="19" ~ 1, Temp=="22"~2, Temp=="25"~0))  %>% pull(temp_shape)

sp_shape_p <- pcomp_pcau  %>% ungroup %>% mutate(temp_shape = case_when(Temp=="19" ~ 1, Temp=="22"~2, Temp=="25"~0))  %>% pull(temp_shape)

comp_fig <- ggplot() +  
     scale_fill_manual(name = "Temperature",labels = temp_name,values=temp_col) +
  # bars for VFL frequencies
   stat_summary(overlay_bottom_cmp, mapping=aes(x=Day, y=Freq*8.5, fill=Temp_f),fun.y = "mean", geom = "bar",position=position_dodge(4), stat="identity", alpha = 0.5, width = 4)+
  # tetra+Glauc
    scale_color_manual(name = "Temperature",labels = temp_name,values=temp_col2) +
    scale_shape_manual(name ="Temperature", labels = temp_name,values=c(16,17,15))+
    scale_linetype_manual(name = "Species",values=c("dotted","solid","dashed"))+
  geom_point(data=pcomp_tetra_glauc,aes(x=Day, y=lnDens,group=Sp_temp,color=Temp_f,shape=Temp_f),position=position_jitterdodge(jitter.width = 1.5, dodge.width = 4),size = 1 ,alpha=0.7,shape=sp_shape_tg)+
    stat_summary(data=pcomp_tetra_glauc,aes(x=Day, y=lnDens,group=Sp_temp,color=Temp_f,linetype = Species), fun.y = "mean", geom = "line",position = position_dodge(4),size=0.6)+
    stat_summary(data=pcomp_tetra_glauc,aes(x=Day, y=lnDens,group=Sp_temp,color=Temp_f),fun.data = mean_se, geom = "errorbar",position = position_dodge(4), size= 0.45, width=3) +
  stat_summary(data=pcomp_tetra_glauc,aes(x=Day, y=lnDens,group=Sp_temp,color=Temp_f,shape=Temp_f),fun.y = "mean", geom = "point",position = position_dodge(4),size = 1.5)+
  # P.caud
  geom_point(data=pcomp_pcau,aes(x=Day, y=lnDens,group=Sp_temp,color=Temp_f),position=position_jitterdodge(jitter.width = 1.5, dodge.width = 4),size = 1 ,alpha=0.7,shape=sp_shape_p)+
    stat_summary(data=pcomp_pcau,aes(x=Day, y=lnDens,group=Sp_temp,color=Temp_f,linetype = Species), fun.y = "mean", geom = "line",position = position_dodge(4),size=0.6)+
    stat_summary(data=pcomp_pcau,aes(x=Day, y=lnDens,group=Sp_temp,color=Temp_f),fun.data = mean_se, geom = "errorbar",position = position_dodge(4), size= 0.45, width=3) +
  stat_summary(data=pcomp_pcau,aes(x=Day, y=lnDens,group=Sp_temp,color=Temp_f,shape=Temp_f),fun.y = "mean", geom = "point",position = position_dodge(4),size = 1.5)+
    facet_wrap(~Trt,scale = "fixed") +
    scale_y_continuous(name = "Protist Ln(Density)", limits = c(0,8.5), sec.axis = sec_axis( trans=~./8.5, name="VFL1-1 Frequency"))+
    scale_x_discrete(limits = c(5,10,15))+
  fig.theme
comp_fig


wt_fsc_cmp <- c0_trim_fscH %>%filter(Strain == "WT-mNG", PredComp=="Yes", Day!=0)

comp_wt_size <- ggplot(data=wt_fsc_cmp,aes(x=Day, y=lgFSC_H, group=Temp_Day,fill=Temp_f,color=Temp_f)) + 
   geom_violin()+ 
    scale_color_manual(name = "Temperature",labels = temp_name,values=temp_col) + 
    scale_fill_manual(name = "Temperature",labels = temp_name,values=temp_col)+
  stat_summary(fun.y = "mean", geom = "point",position = position_dodge(4.5), size=1.5, stroke = 1,shape = 1, color="#9c97a8")+
  stat_summary(fun.y = "mean", geom = "point",position = position_dodge(4.5), size=1.3, shape = 16, color="white")+
    facet_grid(~Trt) +
  theme_bw()+theme(panel.border = element_blank(), panel.grid.major = element_blank(),panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"), strip.background =element_rect(colour="white", fill="#f0f0f0"),plot.title = element_text(color="black", size=14, face="bold",hjust=0.5),axis.ticks.length=unit(-0.15, "cm"), axis.text.x = element_text(margin=unit(c(0.5,0.5,0.5,0.5), "cm")), axis.text.y = element_text(margin=unit(c(0.5,0.5,0.5,0.5), "cm")))
comp_wt_size

fig7 <- ggarrange(comp_fig,comp_wt_size,heights = c(6, 3),labels = c("a", "b"),ncol = 1, nrow = 2,common.legend = TRUE, legend = "bottom")
fig7
```

## fig 8

```{r fig.width=10, fig.height=6}
#merging chlamy density with protists density]
cd2 <- cdenSt %>% rename(Species=Strain)
dens2 <-  rbind(cd2, pd2[,-c("Sp_temp","SpTempTD")]) %>% mutate(Trt=factor(Trt,levels=c("Control","Tetra","Glauc","P.cau","TetraP.cau","GlaucP.cau")),Species=factor(Species,levels=c("VFL1-1","WT-mNG","Tetra","Glauc","P.cau")))

Ap.eco.allSpecies <- ggplot() + 
    scale_color_manual(name = "Species",values=c("#f4ca76","#add0af","#a4a4a4", "#7fb1dd", "#edb494"))+
    geom_point(data=dens2,aes(x=Day, y=lnDens, group=Species,fill=Species,color=Species),position=position_jitterdodge(jitter.width = 1.5, dodge.width = 4),size = 1, alpha=0.5)+
    stat_summary(data=dens2,aes(x=Day, y=lnDens, group=Species,fill=Species,color=Species),fun.y = "mean", geom = "line",position = position_dodge(4),size=0.6,alpha=0.5)+
    stat_summary(data=dens2,aes(x=Day, y=lnDens, group=Species,fill=Species,color=Species),fun.y = "mean", geom = "point",position = position_dodge(4),size = 1)+
    stat_summary(data=dens2,aes(x=Day, y=lnDens,group=Species,fill=Species,color=Species),fun.data = mean_se, geom = "errorbar",position = position_dodge(4),width=3, size= 0.6, alpha=0.5)+
    facet_grid(Temp~Trt,labeller = labeller(Trt = Trt.labs)) +
    fig.theme; Ap.eco.allSpecies
```

# Appendix III Earth Mover's Distance

## fig 1

```{r}
# pulling vfl1-1 frequence over time for each temperature and predator treatment for calculating EMD
for(i in c(19,22,25)){
  for(j in c("Tetra","P.cau","TetraP.cau","Glauc","GlaucP.cau")){
    a <- vfl_sum %>% filter(Temp==i,Trt==j) %>% ungroup%>% dplyr::select(Day,Freq) %>% as.matrix()
    assign(paste0(j,"_",i), a)
  }}

require(emdist)

# calculating EMD
# here we evaluate the relative contribution of Tetrahymena and P.caudatum on prey evoltuion by calculating the similarity of prey evolutionary outcome (vfl1-1 frequency overtime) between 1) Tetrahynena-only treatment vs Tetrahymena+P.caudatum competition treatment 2) P.caudatum-only treatment vs Tetrahymena+P.caudatum competition treatment 3) Tetrahymena-only vs P.caudatum-only treatment across all temperature treatment.
TP_emd <- data_frame(Temp=c("19","22","25"),T_comp = c(emd(Tetra_19,TetraP.cau_19),emd(Tetra_22,TetraP.cau_22),emd(Tetra_25,TetraP.cau_25)),P_comp=c(emd(P.cau_19,TetraP.cau_19),emd(P.cau_22,TetraP.cau_22),emd(P.cau_25,TetraP.cau_25)),T_P=c(emd(Tetra_19,P.cau_19),emd(Tetra_22,P.cau_22),emd(Tetra_25,P.cau_25))) %>% pivot_longer(T_comp:T_P,names_to = "pairs", values_to = "EMD") %>% mutate(pairs=factor(pairs,levels = c("P_comp","T_comp","T_P")))

# Similarly, we evaluate the relative contribution of Glaucoma and P.caudatum on prey evoltuion
GP_emd <- data_frame(Temp=c("19","22","25"),G_comp = c(emd(Glauc_19,GlaucP.cau_19),emd(Glauc_22,GlaucP.cau_22),emd(Glauc_25,GlaucP.cau_25)),P_comp=c(emd(P.cau_19,GlaucP.cau_19),emd(P.cau_22,GlaucP.cau_22),emd(P.cau_25,GlaucP.cau_25)),G_P=c(emd(Glauc_19,P.cau_19),emd(Glauc_22,P.cau_22),emd(Glauc_25,P.cau_25))) %>% pivot_longer(G_comp:G_P,names_to = "pairs", values_to = "EMD") %>% mutate(pairs=factor(pairs,levels = c("P_comp","G_comp","G_P")))

TP_labels <- c("Paramecium:Tetra+Param","Tetrahymena:Tetra+Param","Tetrahymena:Paramecium")
GP_labels <- c("Paramecium:Glauc+Param","Glaucoma:Glauc+Param","Glaucoma:Paramecium")


TP_emd_fig <- ggplot(TP_emd, aes(x=Temp, y=EMD, group=pairs, color = pairs,linetype=pairs))+
    scale_color_manual(name = "Comparison", labels = TP_labels,values=c("#9e9792","#add0af","#eaa064")) + 
    scale_shape_manual(name = "Comparison",labels = TP_labels, values=c(0,1,2)) +
    scale_linetype_manual(name = "Comparison",labels = TP_labels, values=c(1,2,3)) +
  xlab("Temperature")+
    ggtitle("Tetrahymena & Paramecium")+
    geom_point(position=position_dodge(0.5),size =1.5,stroke =1.2,aes(shape=pairs))+
    geom_line(position = position_dodge(0.5))+
  fig.theme

GP_emd_fig  <- ggplot(GP_emd, aes(x=Temp, y=EMD, group=pairs, color = pairs,linetype=pairs))+
    scale_color_manual(name = "Comparison",labels=GP_labels,values=c("#9e9792","#add0af","#eaa064")) +
    scale_shape_manual(name = "Comparison",labels= GP_labels,values=c(0,1,2)) +
    scale_linetype_manual(name = "Comparison",labels = GP_labels, values=c(1,2,3)) +
  xlab("Temperature")+
    ggtitle("Glaucoma & Paramecium")+
    geom_point(position=position_dodge(0.5),size =1.5,stroke =1.2,aes(shape=pairs))+
    geom_line(position = position_dodge(0.5))+
  fig.theme

evo_emd <- ggarrange(TP_emd_fig, GP_emd_fig,labels = c("A", "B"),ncol = 1, nrow =2,common.legend = F)
evo_emd
```

# Appendix IV

## Table 1

ARMA model selection

```{r}
cs_sum <- c0_trim_fscH %>% group_by(Day,Day_f,Block,Jar,Temp,Temp_f,Trt,Rep,Strain) %>% dplyr::summarise(mLgFSC_H = mean(lgFSC_H),medLgFSC_H = median(lgFSC_H),sdLgFSC_H = sd(lgFSC_H),skewLgFSC_H = skewness(lgFSC_H), kurtoLgFSC_H = kurtosis(lgFSC_H))

vs_sng <- cs_sum %>% filter(Strain=="VFL1-1", Trt %in% c("Control","Tetra", "Glauc","P.cau"))
ws_sng <- cs_sum %>% filter(Strain=="WT-mNG", Trt %in% c("Control","Tetra", "Glauc","P.cau"))

# competition data frame
vs_cmp <- cs_sum %>% filter(Strain=="VFL1-1", Trt %in% c("Tetra", "Glauc","TetraP.cau","GlaucP.cau")) %>% mutate(FocalSp=ifelse(Trt %in% c("Tetra","TetraP.cau"), "Tetra", "Glauc"), PredComp=ifelse(Trt %in% c("Tetra","Glauc"), "No","Yes"))
ws_cmp <- cs_sum %>% filter(Strain=="WT-mNG", Trt %in% c("Tetra", "Glauc","TetraP.cau","GlaucP.cau")) %>% mutate(FocalSp=ifelse(Trt %in% c("Tetra","TetraP.cau"), "Tetra", "Glauc"), PredComp=ifelse(Trt %in% c("Tetra","Glauc"), "No","Yes"))

vs_cmp %>% ungroup%>% dplyr::select(Trt,FocalSp,PredComp)
```

```{r}
# WT log density (Ecological dynamics)
lme.wd0 <- lme(lnDens ~ Temp*Trt*Day, data=wt_sng, random = ~ 1|Block) # Lowest. No autocorrelation
lme.wd1 <- lme(lnDens ~ Temp*Trt*Day, data=wt_sng, random = ~ 1|Block,corr = corARMA(form = ~ Day|Block/Trt/Jar,p=1))
lme.wd2 <- lme(lnDens ~ Temp*Trt*Day, data=wt_sng, random = ~ 1|Block,corr = corARMA(form = ~ Day|Block/Trt/Jar,p=2))
lme.wd3 <- lme(lnDens ~ Temp*Trt*Day, data=wt_sng, random = ~ 1|Block,corr = corARMA(form = ~ Day|Block/Trt/Jar,p=3))
AIC(lme.wd0,lme.wd1,lme.wd2,lme.wd3)

# vfl1-1 log density (Ecological dynamics)
lme.vd0 <- lme(lnDens ~ Temp*Trt*Day, data=vfl_sng, random = ~ 1|Block) # Lowest. No autocorrelation
lme.vd1 <- lme(lnDens ~ Temp*Trt*Day, data=vfl_sng, random = ~ 1|Block,corr = corARMA(form = ~ Day|Block/Trt/Jar,p=1))
lme.vd2 <- lme(lnDens ~ Temp*Trt*Day, data=vfl_sng, random = ~ 1|Block,corr = corARMA(form = ~ Day|Block/Trt/Jar,p=2))
lme.vd3 <- lme(lnDens ~ Temp*Trt*Day, data=vfl_sng, random = ~ 1|Block,corr = corARMA(form = ~ Day|Block/Trt/Jar,p=3))
AIC(lme.vd0,lme.vd1,lme.vd2,lme.vd3)

# Freq (Evolutional dynamics)
lme.vf0 <- lme(Freq ~ Temp*Trt*Day, data=vfl_sng, random = ~ 1|Block) # Lowest. No autocorrelation
lme.vf1 <- lme(Freq ~ Temp*Trt*Day, data=vfl_sng, random = ~ 1|Block,corr = corARMA(form = ~ Day|Block/Trt/Jar,p=1))
lme.vf2 <- lme(Freq ~ Temp*Trt*Day, data=vfl_sng, random = ~ 1|Block,corr = corARMA(form = ~ Day|Block/Trt/Jar,p=2))
lme.vf3 <- lme(Freq ~ Temp*Trt*Day, data=vfl_sng, random = ~ 1|Block,corr = corARMA(form = ~ Day|Block/Trt/Jar,p=3))
AIC(lme.vf0,lme.vf1,lme.vf2,lme.vf3)

# Chlamydomonas wt size (Phenotypic dynamics)
lme.wph0 <- lme(mLgFSC_H ~ Temp*Trt*Day, data= vs_sng, random = ~ 1|Block) # Lowest. No autocorrelation
lme.wph1 <- lme(mLgFSC_H ~ Temp*Trt*Day, data= vs_sng, random = ~ 1|Block,corr = corARMA(form = ~ Day|Block/Trt/Jar,p=1))
lme.wph2 <- lme(mLgFSC_H ~ Temp*Trt*Day, data= vs_sng, random = ~ 1|Block,corr = corARMA(form = ~ Day|Block/Trt/Jar,p=2))
lme.wph3 <- lme(mLgFSC_H ~ Temp*Trt*Day, data= vs_sng, random = ~ 1|Block,corr = corARMA(form = ~ Day|Block/Trt/Jar,p=3))

AIC(lme.wph0,lme.wph1,lme.wph2,lme.wph3)
```

## Table 2

Temperautre and predator identity effects on C.reinhardtii vfl1-1 clonal frequencies

```{r}
lmm_vf <- lmer(Freq ~ Temp*Trt*Day+(1|Block), data=vfl_sng) 
summary(lmm_vf); Anova(lmm_vf)
```

## Table 3

Temperautre and predator identity effects on Mean size of vfl1-1

```{r}
lmm_vs_av <- lmer(mLgFSC_H ~ Temp*Trt*Day+(1|Block), data=vs_sng) 
summary(lmm_vs_av);Anova(lmm_vs_av)
```

## Table 4

```{r}
lmm_ws_av <- lmer(mLgFSC_H ~ Temp*Trt*Day+(1|Block), data=ws_sng) 
summary(lmm_ws_av);Anova(lmm_ws_av)
```

## Table 5

```{r}
vf_dltT.s <- lmer(dlt_Freq_T ~ Temp*Trt*Day+(1|Block), data=vdltf_sng) # everything significant!
summary(vf_dltT.s);Anova(vf_dltT.s)
```

## Table 6

```{r}
require(MASS)
c.full.r<-lm(r~Temp*Trt*Strain, data =c_r)
c_r.step <- stepAIC(c.full.r, direction = "both", trace = FALSE)
summary(c_r.step)

c.full.max<-lm(lnDens~Temp*Trt*Strain, data =c_max)
c_K.step <- stepAIC(c.full.max, direction = "both", trace = FALSE)
summary(c_K.step)
```

## Table 7

```{r}
# protists r and K linear model 
full.r<-lm(r~Temp*Species*PredComp, data =pd_r)
p_r.step <- stepAIC(full.r, direction = "both", trace = FALSE)
summary(p_r.step )


full.md<-lm(lnDens~Temp*Species*PredComp, data =pd_max)
p_K.step <- stepAIC(full.md, direction = "both", trace = FALSE)
summary(p_K.step)
```

## Table 8

```{r}
# medium 
lmm_w_med <- lmer(medLgFSC_H ~ Temp*Trt*Day+(1|Block), data=ws_sng); Anova(lmm_w_med)

# sd
lmm_w_sd <- lmer(sdLgFSC_H ~ Temp*Trt*Day+(1|Block), data=ws_sng); Anova(lmm_w_sd)

# skewness
lmm_w_sk <- lmer(skewLgFSC_H ~ Temp*Trt*Day+(1|Block), data=ws_sng); Anova(lmm_w_sk)

# Kurtosis
lmm_w_kt <- lmer(kurtoLgFSC_H ~ Temp*Trt*Day+(1|Block), data=ws_sng); Anova(lmm_w_kt)
```

## Table 9

temperautre and predator competition on C.reinhardtii vfl1-1 clonal frequencies

```{r}
lmm_vf.c <- lmer(Freq ~ Temp*FocalSp*PredComp*Day+(1|Block), data=vfl_cmp) 
summary(lmm_vf.c); Anova(lmm_vf.c)
```

## Table 10

Delta Temp

```{r}
v19_mean_cmp <- vfl_cmp %>% filter(Temp == 19 ) %>% group_by(Temp,Trt,Day) %>% summarize(m19Freq=mean(Freq)) %>% ungroup %>% dplyr::select(-Temp)
vdltf_cmp <- vfl_cmp %>% filter( !Temp == 19 ) %>% left_join(.,v19_mean_cmp) %>% mutate(dlt_Freq_T=Freq-m19Freq)

vf_dltT.c <- lmer(dlt_Freq_T  ~ Temp*FocalSp*PredComp*Day+(1|Block), data=vdltf_cmp) 
summary(vf_dltT.c);Anova(vf_dltT.c)
```

## Table 11

```{r}
lmm_wscmp_av <- lmer(mLgFSC_H ~ Temp*FocalSp*PredComp*Day+(1|Block), data=ws_cmp) 
summary(lmm_wscmp_av);Anova(lmm_wscmp_av)
```

## Table 12

```{r}
lmm_vscmp_av <- lmer(mLgFSC_H ~ Temp*FocalSp*PredComp*Day+(1|Block), data=vs_cmp) 
summary(lmm_vscmp_av);Anova(lmm_vscmp_av)
```
